---
title: "Variable Overview — CoLaus / OsteoLaus Study"
subtitle: "Data Cleaning Guidelines & Reproducibility Reference"
author: ""
date: last-modified
date-format: "DD MMMM YYYY"

format:
  html:
    toc: false
    toc-depth: 3
    toc-location: left
    toc-title: "Contents"
    number-sections: true
    theme: cosmo
    code-fold: true
    code-tools: true
    embed-resources: true          # single self-contained HTML file
    smooth-scroll: true
    df-print: paged
    page-layout: full
  
execute:
  echo: true
  warning: false
  message: false
  cache: false
---

```{r}
#| label: setup
#| include: false

library(tidyverse)    # data wrangling & ggplot2
library(DT)           # interactive tables
library(reactable)    # rich interactive tables with grouping
library(gt)           # publication-quality static tables
library(htmltools)    # HTML helpers for callout rendering
library(knitr)        # kable for simple tables
```

```{r}
#| label: colour-palette
#| include: false

# Consistent colour coding used throughout this document
ROLE_COLOURS <- c(
  "Primary Key"              = "#1A3A5C",
  "Temporal Key"             = "#2E86AB",
  "Primary Outcome"          = "#E84855",
  "Secondary Outcome"        = "#F4845F",
  "Exposure"                 = "#6A4C93",
  "Fixed Covariate"          = "#2D6A4F",
  "Time-Varying Covariate"   = "#52B788",
  "Source"                   = "#B7B7B7",
  "Validation"               = "#F4A261",
  "Inactive"                 = "#DCDCDC",
  "Excluded"                 = "#F2F2F2"
)

STATUS_COLOURS <- c(
  "Complete"     = "#2D6A4F",
  "Open ⚠"       = "#E84855",
  "Partial"      = "#F4A261"
)
```

# Study Design & Data Structure

## Cohorts

CoLaus is a population-based cohort of adults from Lausanne, Switzerland. It has five measurement time points: Baseline, F1, F2, F3, and F4. Individual CSV files are provided per time point. FFQ (food frequency questionnaire) data are available from F1 onwards.

OsteoLaus is a sub-cohort of CoLaus including female participants only. It has five visit time points: Baseline, V2, V3, V4, and V5. OsteoLaus visits are mapped to the closest CoLaus time point (many-to-one). Because OsteoLaus contains only female participants, the `sex` variable is excluded from analysis.

## Merge Logic

```         
CoLaus (Baseline–F4)
        │
        │  merge on: pt  (1:1 within wave)
        │
OsteoLaus (Baseline–V5) ──── mapped to nearest CoLaus wave ──── exam_date_iso (join key)
```

-   The `pt` variable is the **primary key** (participant-level unique identifier).\
-   `exam_date_iso` (ISO 8601 date) is the **temporal join key**.\
-   OsteoLaus → CoLaus mapping is **many-to-one**: multiple OsteoLaus visits may link to one CoLaus wave.

# Role Taxonomy & Colour Key

```{r}
#| label: role-legend
#| echo: false

role_legend <- tibble(
  Role = names(ROLE_COLOURS),
  `Colour` = ROLE_COLOURS,
  Description = c(
    "Unique participant identifier. Must never be missing.",
    "Examination date used for dataset joins and time calculations.",
    "Main study outcome.",
    "Outcome used in secondary or sensitivity analyses.",
    "Primary dietary exposure under investigation.",
    "Baseline-only covariate; assumed stable over time.",
    "Covariate re-measured at each time point.",
    "Raw variable used solely to derive another variable; not used directly in models.",
    "Variable used to cross-validate another variable.",
    "Present in dataset but not used in this analysis.",
    "Dropped before analysis; not carried forward."
  )
)

role_legend |>
  select(Role, Description) |>
  gt() |>
  tab_style(
    style = list(cell_fill(color = "#1A3A5C"), cell_text(color = "white", weight = "bold")),
    locations = cells_body(columns = Role, rows = Role == "Primary Key")
  ) |>
  tab_style(
    style = list(cell_fill(color = "#2E86AB"), cell_text(color = "white")),
    locations = cells_body(columns = Role, rows = Role == "Temporal Key")
  ) |>
  tab_style(
    style = list(cell_fill(color = "#E84855"), cell_text(color = "white", weight = "bold")),
    locations = cells_body(columns = Role, rows = Role == "Primary Outcome")
  ) |>
  tab_style(
    style = list(cell_fill(color = "#F4845F"), cell_text(color = "white")),
    locations = cells_body(columns = Role, rows = Role == "Secondary Outcome")
  ) |>
  tab_style(
    style = list(cell_fill(color = "#6A4C93"), cell_text(color = "white", weight = "bold")),
    locations = cells_body(columns = Role, rows = Role == "Exposure")
  ) |>
  tab_style(
    style = list(cell_fill(color = "#2D6A4F"), cell_text(color = "white")),
    locations = cells_body(columns = Role, rows = Role == "Fixed Covariate")
  ) |>
  tab_style(
    style = list(cell_fill(color = "#52B788"), cell_text(color = "white")),
    locations = cells_body(columns = Role, rows = Role == "Time-Varying Covariate")
  ) |>
  tab_options(table.width = pct(100))
```

**Multiplicity notation:**

| Code | Meaning |
|-----------------------------|-------------------------------------------|
| `1..1` | Exactly one value per participant (never missing) |
| `0..1` | At most one value per participant (may be missing) |
| `1..*` | One or more values across waves (must have at least one) |
| `0..*` | Zero or more values across waves (longitudinal; may be missing at some waves) |

# Variable Definitions {#sec-variables}

```{r}
#| label: variable-data
#| include: false

# ─── Master variable table ────────────────────────────────────────────────────
# Columns:
#   source       : "CoLaus" | "OsteoLaus" | "Both"
#   section      : Thematic section label
#   original     : Variable name in raw dataset
#   derived      : Derived variable name (NA if none)
#   data_type    : Storage type
#   role         : Analytical role (see taxonomy)
#   multiplicity : UML multiplicity
#   levels       : Factor levels or units
#   notes        : Description / derivation logic
#   missing      : Missing value handling rule
#   open         : TRUE if item requires investigator decision

vars <- tribble(
  ~source,~section, ~original,~derived, ~data_type, ~role, ~multiplicity, ~levels, ~notes,  ~missing, ~open,

  # ── Identifiers
  "Both",     "Identifiers & Keys",         "pt",                  NA,                          "Integer",        "Primary Key",              "1..1",        "—",                                                                                                        "Unique participant identifier for the CoLaus/OsteoLaus cohort.",                                                                                                         "Exclude participant if not recorded.",                                                                                     FALSE,
  "Both",     "Identifiers & Keys",         "datexam",             "exam_date_iso",             "Numeric date",   "Source",                   "1..*",        "Raw: DDMonYYYY",                                                                                           "Examination date. Converted to ISO 8601 (YYYY-MM-DD).",                                                                                                                  "—",                                                                                                                        FALSE,
  "Both",     "Identifiers & Keys",         "exam_date_iso",       NA,                          "Date",           "Temporal Key",             "1..*",        "ISO 8601 (YYYY-MM-DD)",                                                                                    "Join key. Many-to-one mapping from OsteoLaus visits to closest CoLaus wave.",                                                                                            "Exclude participant if not recorded.",                                                                                     FALSE,
  "CoLaus",   "Identifiers & Keys",         "age",                 NA,                          "Numeric (float)","Time-Varying Covariate",   "1..*",        "Years",                                                                                                    "Age calculated at each exam_date_iso.",                                                                                                                                  "Exclude at baseline if missing. Back-fill using exam_date_iso for follow-ups.",                                            FALSE,

  # ── Excluded
  "CoLaus",   "Excluded Variables",         "sex",                 NA,                          "Factor",         "Excluded",                 "1..1",        "1 = Male; 0 = Female",                                                                                     "Excluded: OsteoLaus contains only female participants.",                                                                                                                 "N/A (excluded)",                                                                                                           FALSE,
  "CoLaus",   "Excluded Variables",         "ethori_self",         NA,                          "Factor",         "Excluded",                 "0..1",        "A=Asian; B=Black/African; W=White; O=Other; X=Unknown",                                                    "Excluded: not used as covariate.",                                                                                                                                       "N/A (excluded)",                                                                                                           FALSE,
  "CoLaus",   "Excluded Variables",         "mrtsts2",             NA,                          "Factor",         "Excluded",                 "0..*",        "0 = Living alone; 1 = Living in couple",                                                                   "Marital status. Excluded: not used as covariate.",                                                                                                                       "N/A (excluded)",                                                                                                           FALSE,

  # ── Education
  "CoLaus",   "Fixed Covariates",           "edtyp4",              "education_level",           "Integer",        "Source",                   "0..1",        "1=University; 2=High school; 3=Apprenticeship; 4=Mandatory",                                               "ISCED mapping: University=6–8; High school (Gymnasium/Matura)=3; Apprenticeship (VET)=3–4; Mandatory=1–2.",                                                              "—",                                                                                                                        FALSE,
  "CoLaus",   "Fixed Covariates",           "education_level",     NA,                          "Factor",         "Fixed Covariate",          "0..1",        "1=Low (ISCED 0–2); 2=Medium (ISCED 3–4); 3=High (ISCED 5–8)",                                              "Three-level education variable derived from edtyp4.",                                                                                                                    "Exclude participant if not recorded.",                                                                                     FALSE,

  # ── Alcohol
  "CoLaus",   "Alcohol",                    "alcuse",              NA,                          "Binary",         "Validation",               "0..*",        "0 = No; 1 = Yes",                                                                                          "Used to cross-validate non-drinker classification in derived variable.",                                                                                                 "—",                                                                                                                        FALSE,
  "CoLaus",   "Alcohol",                    "alcool4",             NA,                          "Binary",         "Validation",               "0..*",        "0 = Non-drinker; 1 = Drinker",                                                                             "Redundant if alcuse is present. Check data completeness between the two.",                                                                                               "—",                                                                                                                        FALSE,
  "CoLaus",   "Alcohol",                    "conso_hebdo",         "alcohol_category",          "Numeric",        "Source",                   "1..*",        "Units/week (1 unit = 10–12 g)",                                                                            "Primary source for Baseline where sumalco is missing by design.",                                                                                                        "See alcohol_category.",                                                                                                    FALSE,
  "CoLaus",   "Alcohol",                    "sumalco",             "alcohol_category",          "Numeric",        "Source",                   "0..*",        "g/day",                                                                                                    "Missing by design at Baseline. Primary input from F1 onwards.",                                                                                                          "See alcohol_category.",                                                                                                    FALSE,
  "CoLaus",   "Alcohol",                    "alcohol_category",    NA,                          "Factor",         "Time-Varying Covariate",   "0..*",        "1=Non-drinker (0 g/day); 2=Light (>0–6); 3=Moderate (>6–12); 4=Heavy (>12)",                               "Baseline: derived from conso_hebdo. F1+: sumalco preferred; conso_hebdo used if sumalco missing.",                                                                        "⚠ OPEN: If both sources missing at a wave, carry forward (or backward) from adjacent wave. Exclude if no category can be imputed. Confirm proportion changing category over time.", TRUE,

  # ── Smoking
  "CoLaus",   "Smoking",                    "sbsmk",               NA,                          "Factor",         "Time-Varying Covariate",   "0..*",        "0=Never; 1=Former; 2=Current",                                                                             "Smoking history screening question.",                                                                                                                                    "⚠ OPEN: Exclude if missing at baseline. Carry baseline status forward for missing follow-ups. Confirm proportion changing category over time.",                             TRUE,

  # ── Diabetes
  "CoLaus",   "Diabetes",                   "dbtld",               "diabetes_status",           "Factor",         "Source",                   "0..*",        "Baseline/F1/F3: 0=No; 1=Yes; 9=DK. F2: 0=No; 1=Yes; 9=No data",                                           "Self-reported diagnosis ('Ever told by doctor').",                                                                                                                       "—",                                                                                                                        FALSE,
  "CoLaus",   "Diabetes",                   "agediag_dbts",        NA,                          "Numeric",        "Excluded",                 "0..1",        "—",                                                                                                        "Age at diabetes diagnosis. Could distinguish Type 1 vs 2 but mostly missing. Excluded.",                                                                                "—",                                                                                                                        FALSE,
  "CoLaus",   "Diabetes",                   "dbdrg",               "diabetes_status",           "Factor",         "Source",                   "0..*",        "Baseline/F1/F3: 0=No; 1=Yes; 9=DK. F2: 0=No; 1=Yes; 8=N/A; 9=No data",                                    "Self-reported history of diabetes treatment.",                                                                                                                           "—",                                                                                                                        FALSE,
  "CoLaus",   "Diabetes",                   "orldrg",              "diabetes_status",           "Factor",         "Source",                   "0..*",        "0=No; 1=Yes; 8=N/A; 9=No data",                                                                            "Use of oral antidiabetic agents.",                                                                                                                                       "⚠ OPEN: Verify whether missing values originate from non-diabetics (expected) or data error.",                             TRUE,
  "CoLaus",   "Diabetes",                   "insn",                "diabetes_status",           "Factor",         "Source",                   "0..*",        "0=No; 1=Yes; 8=N/A; 9=No data",                                                                            "Insulin use. Critical for Level 3 classification.",                                                                                                                      "—",                                                                                                                        FALSE,
  "CoLaus",   "Diabetes",                   "antiDIAB",            "diabetes_status",           "Binary",         "Source",                   "0..*",        "0=No; 1=Yes",                                                                                              "General antidiabetic drug flag.",                                                                                                                                        "—",                                                                                                                        FALSE,
  "CoLaus",   "Diabetes",                   "DIAB",                "diabetes_status",           "Binary",         "Source",                   "0..*",        "0=No; 1=Yes",                                                                                              "Fasting plasma glucose ≥ 7.0 mmol/L. Primary clinical marker.",                                                                                                         "—",                                                                                                                        FALSE,
  "CoLaus",   "Diabetes",                   "DIAB_Hb",             "diabetes_status",           "Binary",         "Source",                   "0..*",        "0=No; 1=Yes",                                                                                              "HbA1c ≥ 48 mmol/mol. Missing by design; available from F2 onwards.",                                                                                                    "—",                                                                                                                        FALSE,
  "CoLaus",   "Diabetes",                   "DIAB2",               "diabetes_status",           "Factor",         "Validation",               "0..*",        "Baseline/F2/F3: 0=Normal; 1=IFG; 2=Diabetes. F1: 0=No; 1=Yes",                                             "Combined IGT/Diabetes flag for prediabetes verification.",                                                                                                               "⚠ OPEN: Resolve handling of simplified binary coding at F1 vs 3-level coding at other waves.",                             TRUE,
  "CoLaus",   "Diabetes",                   "metab_synd",          NA,                          "Binary",         "Inactive",                 "0..*",        "0=No; 1=Yes",                                                                                              "Metabolic Syndrome (ATP-III). Redundant if using specific diabetes_status.",                                                                                             "—",                                                                                                                        FALSE,
  "CoLaus",   "Diabetes",                   "diabetes_status",     NA,                          "Factor",         "Time-Varying Covariate",   "0..*",        "1=No diabetes; 2=Diabetes (oral/lab); 3=Diabetes (+insulin); 4=IGT/Prediabetes",                            "Derived from: dbtld, dbdrg, orldrg, insn, antiDIAB, DIAB, DIAB_Hb, DIAB2.",                                                                                             "—",                                                                                                                        FALSE,

  # ── Medications
  "CoLaus",   "Medications (ATC)",          "ATC1-21",             "multiple drug flags",       "Character",      "Source",                   "0..*",        "—",                                                                                                        "Raw ATC codes for all reported medications. Used to derive all drug status flags below.",                                                                                "—",                                                                                                                        FALSE,
  "CoLaus",   "Medications (ATC)",          "ATC_OTC1-17",         NA,                          "Character",      "Source",                   "0..*",        "—",                                                                                                        "Over-the-counter drug ATC codes.",                                                                                                                                       "—",                                                                                                                        FALSE,
  "CoLaus",   "Medications (ATC)",          "Diet_compl",          NA,                          "Numeric",        "Inactive",                 "0..*",        "—",                                                                                                        "Total number of dietary complements reported.",                                                                                                                          "—",                                                                                                                        FALSE,
  "CoLaus",   "Medications (ATC)",          "hypolip_drug_status", "lip_status",               "Binary",         "Time-Varying Covariate",   "0..*",        "0=No; 1=Yes",                                                                                              "Any ATC code starting with C10. Used to verify hypolip or hctld.",                                                                                                       "—",                                                                                                                        FALSE,
  "CoLaus",   "Medications (ATC)",          "corticoids_status",   NA,                          "Binary",         "Time-Varying Covariate",   "0..*",        "0=No; 1=Yes",                                                                                              "Any ATC code starting with H02.",                                                                                                                                        "—",                                                                                                                        FALSE,
  "CoLaus",   "Medications (ATC)",          "vitD_status",         NA,                          "Binary",         "Time-Varying Covariate",   "0..*",        "0=No; 1=Yes",                                                                                              "Any ATC code starting with A11.",                                                                                                                                        "—",                                                                                                                        FALSE,
  "CoLaus",   "Medications (ATC)",          "calcium_status",      NA,                          "Binary",         "Time-Varying Covariate",   "0..*",        "0=No; 1=Yes",                                                                                              "⚠ OPEN: ATC prefix A12A (incl. A12AX) — to be confirmed.",                                                                                                               "—",                                                                                                                        TRUE,
  "CoLaus",   "Medications (ATC)",          "benzo_status",        NA,                          "Binary",         "Time-Varying Covariate",   "0..*",        "0=No; 1=Yes",                                                                                              "⚠ OPEN: ATC prefix for benzodiazepines to be specified.",                                                                                                                "—",                                                                                                                        TRUE,

  # ── Cardiovascular
  "CoLaus",   "Cardiovascular History",     "miac",                "cdv_event",                "Factor",         "Source",                   "0..*",        "0=No; 1=Yes; 9=DK",                                                                                        "Myocardial infarction history.",                                                                                                                                         "If any CV flag = 1, code cdv_event = 1. Exclude only if all CV flags are 0 or all missing.",                               FALSE,
  "CoLaus",   "Cardiovascular History",     "strk",                "cdv_event",                "Factor",         "Source",                   "0..*",        "0=No; 1=Yes; 9=DK",                                                                                        "Stroke history.",                                                                                                                                                        "Same as miac.",                                                                                                            FALSE,
  "CoLaus",   "Cardiovascular History",     "chf",                 "cdv_event",                "Factor",         "Source",                   "0..*",        "0=No; 1=Yes; 9=DK",                                                                                        "Congestive heart failure history.",                                                                                                                                      "Same as miac.",                                                                                                            FALSE,
  "CoLaus",   "Cardiovascular History",     "cad",                 "cdv_event",                "Factor",         "Source",                   "0..*",        "0=No; 1=Yes; 9=DK",                                                                                        "Coronary artery disease history.",                                                                                                                                       "Same as miac.",                                                                                                            FALSE,
  "CoLaus",   "Cardiovascular History",     "angn",                "cdv_event",                "Factor",         "Source",                   "0..*",        "0=No; 1=Yes; 9=DK",                                                                                        "Angina history.",                                                                                                                                                        "Same as miac.",                                                                                                            FALSE,
  "CoLaus",   "Cardiovascular History",     "cmp",                 "cdv_event",                "Factor",         "Source",                   "0..*",        "0=No; 1=Yes; 9=DK",                                                                                        "Cardiomyopathy history.",                                                                                                                                                "Same as miac.",                                                                                                            FALSE,
  "CoLaus",   "Cardiovascular History",     "hdc",                 "cdv_event",                "Factor",         "Source",                   "0..*",        "0=No; 1=Yes; 9=DK",                                                                                        "Hypertensive heart disease.",                                                                                                                                            "Same as miac.",                                                                                                            FALSE,
  "CoLaus",   "Cardiovascular History",     "hdv",                 "cdv_event",                "Factor",         "Source",                   "0..*",        "0=No; 1=Yes; 9=DK",                                                                                        "Heart valve disease.",                                                                                                                                                   "Same as miac.",                                                                                                            FALSE,
  "CoLaus",   "Cardiovascular History",     "artm",                "cdv_event",                "Factor",         "Source",                   "0..*",        "0=No; 1=Yes; 9=DK",                                                                                        "Arrhythmia.",                                                                                                                                                            "Same as miac.",                                                                                                            FALSE,
  "CoLaus",   "Cardiovascular History",     "vslg",                "cdv_event",                "Factor",         "Source",                   "0..*",        "0=No; 1=Yes; 9=DK",                                                                                        "Vascular surgery history.",                                                                                                                                              "Same as miac.",                                                                                                            FALSE,
  "CoLaus",   "Cardiovascular History",     "ccth",                "cdv_event",                "Factor",         "Source",                   "0..*",        "0=No; 1=Yes; 9=DK",                                                                                        "Cardiac catheterisation history.",                                                                                                                                       "Same as miac.",                                                                                                            FALSE,
  "CoLaus",   "Cardiovascular History",     "cabg",                "cdv_event",                "Factor",         "Source",                   "0..*",        "0=No; 1=Yes; 9=DK",                                                                                        "Bypass surgery (CABG) history.",                                                                                                                                         "Same as miac.",                                                                                                            FALSE,
  "CoLaus",   "Cardiovascular History",     "pcin",                "cdv_event",                "Factor",         "Source",                   "0..*",        "0=No; 1=Yes; 9=DK",                                                                                        "Stent (PCI) history.",                                                                                                                                                   "Same as miac.",                                                                                                            FALSE,
  "CoLaus",   "Cardiovascular History",     "agediag_cv (multiple)", NA,                       "Numeric",        "Inactive",                 "0..*",        "—",                                                                                                        "Ages at each individual CVD diagnosis. Not used in main analysis.",                                                                                                      "N/A (inactive)",                                                                                                           FALSE,
  "CoLaus",   "Cardiovascular History",     "cvdbase_adj",         NA,                          "Binary",         "Validation",               "0..*",        "1=Yes",                                                                                                    "⚠ OPEN: Definition unclear — requires clarification with data custodian.",                                                                                               "—",                                                                                                                        TRUE,
  "CoLaus",   "Cardiovascular History",     "cdv_event",           NA,                          "Factor",         "Time-Varying Covariate",   "0..*",        "0=No; 1=Yes",                                                                                              "Composite CVD event: Yes if any individual CV flag is positive.",                                                                                                        "cdv_event = 1 if any component = 1. Exclude if all components are 0 or all missing.",                                     FALSE,

  # ── Blood pressure / lipids
  "CoLaus",   "Blood Pressure & Lipids",    "crbpmed",             NA,                          "Factor",         "Source",                   "0..*",        "0=No; 1=Yes; 8=N/A; 9=DK",                                                                                "Self-report: 'Do you take medication for high blood pressure?'",                                                                                                         "⚠ OPEN: Missing value strategy to be defined.",                                                                            TRUE,
  "CoLaus",   "Blood Pressure & Lipids",    "antiHTA",             NA,                          "Binary",         "Source",                   "0..*",        "0=No; 1=Yes",                                                                                              "Documented antihypertensive drug treatment.",                                                                                                                            "⚠ OPEN: Missing value strategy to be defined.",                                                                            TRUE,
  "CoLaus",   "Blood Pressure & Lipids",    "HTA",                 NA,                          "Binary",         "Source",                   "0..*",        "0=No; 1=Yes",                                                                                              "Hypertension: Yes if BP ≥ 140/90 mmHg OR on antiHTA.",                                                                                                                  "⚠ OPEN: Missing value strategy to be defined.",                                                                            TRUE,
  "CoLaus",   "Blood Pressure & Lipids",    "hta_status",          NA,                          "Factor",         "Inactive",                 "0..*",        "—",                                                                                                        "Derived hypertension status variable.",                                                                                                                                  "⚠ OPEN: Missing value strategy to be defined.",                                                                            TRUE,
  "CoLaus",   "Blood Pressure & Lipids",    "hctld",               "lip_status",               "Factor",         "Source",                   "0..*",        "Baseline/F1/F3: 0=No; 1=Yes; 9=DK. F2: 0=No; 1=Yes; 9=No data",                                           "Self-report: 'Ever told you had high cholesterol.'",                                                                                                                     "⚠ OPEN: Missing value strategy to be defined.",                                                                            TRUE,
  "CoLaus",   "Blood Pressure & Lipids",    "hypolip",             "lip_status",               "Binary",         "Source",                   "0..*",        "0=No; 1=Yes",                                                                                              "Hypolipidemic treatment. Missing by design at Baseline; F1 onwards.",                                                                                                   "⚠ OPEN: Missing value strategy to be defined.",                                                                            TRUE,
  "CoLaus",   "Blood Pressure & Lipids",    "lip_status",          NA,                          "Factor",         "Inactive",                 "0..*",        "—",                                                                                                        "Derived lipid/statin status.",                                                                                                                                           "⚠ OPEN: Missing value strategy to be defined.",                                                                            TRUE,

  # ── Hormonal
  "CoLaus",   "Hormonal / Reproductive",    "esthrp",              "hrt_status",               "Factor",         "Source",                   "0..*",        "0=No; 1=Yes; 9=DK",                                                                                        "Self-report: 'Ever taken HRT?' (Baseline screening).",                                                                                                                   "—",                                                                                                                        FALSE,
  "CoLaus",   "Hormonal / Reproductive",    "esthrpage",           "hrt_status",               "Numeric",        "Source",                   "0..*",        "99=DK",                                                                                                    "Age at HRT start. Used to verify whether HRT was active at Baseline.",                                                                                                  "—",                                                                                                                        FALSE,
  "CoLaus",   "Hormonal / Reproductive",    "bthc",                "hrt_status",               "Binary",         "Source",                   "0..*",        "—",                                                                                                        "F1 onwards. Current HRT use at time of visit.",                                                                                                                          "—",                                                                                                                        FALSE,
  "CoLaus",   "Hormonal / Reproductive",    "subsp",               "hrt_status",               "Numeric",        "Source",                   "0..*",        "—",                                                                                                        "F1 onwards. Age at HRT cessation.",                                                                                                                                      "—",                                                                                                                        FALSE,
  "CoLaus",   "Hormonal / Reproductive",    "subyr",               "hrt_status",               "Numeric",        "Source",                   "0..*",        "—",                                                                                                        "F1 onwards. Cumulative duration of HRT exposure.",                                                                                                                       "—",                                                                                                                        FALSE,
  "CoLaus",   "Hormonal / Reproductive",    "hrt_status",          NA,                          "Factor",         "Time-Varying Covariate",   "1..*",        "Never; Former; Current",                                                                                   "Back-filled for Baseline visit.",                                                                                                                                        "—",                                                                                                                        FALSE,
  "CoLaus",   "Hormonal / Reproductive",    "povdage",             "pcos_status",              "Numeric",        "Inactive",                 "0..*",        "—",                                                                                                        "F1 onwards. Age at PCOS/POVD diagnosis.",                                                                                                                                "—",                                                                                                                        FALSE,
  "CoLaus",   "Hormonal / Reproductive",    "pcos_status",         NA,                          "Binary",         "Inactive",                 "0..*",        "—",                                                                                                        "Yes if povdage is non-null. Proxy for androgen/estrogen imbalance.",                                                                                                     "—",                                                                                                                        FALSE,

  # ── Muscle strength (CoLaus)
  "CoLaus",   "Muscle Strength (CoLaus)",   "handgrip",            "ewgsop2_sarcopenia_stage", "Numeric (float)","Primary Outcome",          "1..*",        "kg",                                                                                                       "Peak handgrip force. Dominant-hand measurement.",                                                                                                                        "⚠ OPEN: Missing value handling to be defined.",                                                                            TRUE,
  "CoLaus",   "Muscle Strength (CoLaus)",   "lateralite",          NA,                          "Factor",         "Inactive",                 "0..*",        "1=Right; 2=Left; 3=Ambidextrous",                                                                          "Hand preference; used to determine dominant hand.",                                                                                                                      "N/A (inactive)",                                                                                                           FALSE,
  "CoLaus",   "Muscle Strength (CoLaus)",   "handgrip_com",        NA,                          "Factor",         "Inactive",                 "0..*",        "Baseline: 0=No; 1=Yes. F2/F3: 0=None; 1=Pain/arthrosis; 2=No time/rejected",                               "Qualitative notes on grip test. Not available for F1.",                                                                                                                  "N/A (inactive)",                                                                                                           FALSE,

  # ── Anthropometrics
  "CoLaus",   "Anthropometrics",            "ht",                  "BMI",                      "Numeric (float)","Time-Varying Covariate",   "0..*",        "cm",                                                                                                       "Height in cm.",                                                                                                                                                          "Carry forward/backward from any available wave (expected stable).",                                                        FALSE,
  "CoLaus",   "Anthropometrics",            "wt",                  "BMI",                      "Numeric (float)","Source",                   "0..*",        "kg",                                                                                                       "Weight in kg. Not available for F2/F3.",                                                                                                                                 "⚠ OPEN: Confirm whether BMI is available for F2/F3 despite weight being missing.",                                         TRUE,
  "CoLaus",   "Anthropometrics",            "BMI",                 "BMI_category",             "Numeric (float)","Time-Varying Covariate",   "0..*",        "kg/m²",                                                                                                    "Body Mass Index = weight / height².",                                                                                                                                    "—",                                                                                                                        FALSE,
  "CoLaus",   "Anthropometrics",            "BMI_category",        NA,                          "Factor",         "Time-Varying Covariate",   "0..*",        "1=Underweight; 2=Normal (ref); 3=Overweight; 4=Obese",                                                      "Four-level BMI classification.",                                                                                                                                         "—",                                                                                                                        FALSE,
  "CoLaus",   "Anthropometrics",            "WHR",                 NA,                          "Numeric (float)","Inactive",                "0..*",        "—",                                                                                                        "Waist-to-hip ratio; marker of central adiposity. Not used in main analysis.",                                                                                            "—",                                                                                                                        FALSE,
  "CoLaus",   "Anthropometrics",            "bmpsc",               NA,                          "Numeric (float)","Time-Varying Covariate",  "0..*",        "% fat mass",                                                                                               "Bioimpedance Analysis (BIA) % fat mass.",                                                                                                                                "—",                                                                                                                        FALSE,

  # ── Physical activity
  "CoLaus",   "Physical Activity",          "PAFQ_SE",             "pa_levels",                "Numeric",        "Source",                   "0..*",        "min/day",                                                                                                  "Sedentary time.",                                                                                                                                                        "—",                                                                                                                        FALSE,
  "CoLaus",   "Physical Activity",          "PAFQ_SE_pct",         NA,                          "Numeric",        "Source",                   "0..*",        "% of day",                                                                                                 "% daily time spent sedentary.",                                                                                                                                          "—",                                                                                                                        FALSE,
  "CoLaus",   "Physical Activity",          "PAFQ_LPA",            "pa_levels",                "Numeric",        "Source",                   "0..*",        "min/day",                                                                                                  "Light physical activity.",                                                                                                                                               "—",                                                                                                                        FALSE,
  "CoLaus",   "Physical Activity",          "PAFQ_LPA_pct",        NA,                          "Numeric",        "Source",                   "0..*",        "% of day",                                                                                                 "% daily time in light PA.",                                                                                                                                              "—",                                                                                                                        FALSE,
  "CoLaus",   "Physical Activity",          "PAFQ_MPA",            "pa_levels",                "Numeric",        "Source",                   "0..*",        "min/day",                                                                                                  "Moderate physical activity.",                                                                                                                                            "—",                                                                                                                        FALSE,
  "CoLaus",   "Physical Activity",          "PAFQ_MPA_pct",        NA,                          "Numeric",        "Source",                   "0..*",        "% of day",                                                                                                 "% daily time in moderate PA.",                                                                                                                                           "—",                                                                                                                        FALSE,
  "CoLaus",   "Physical Activity",          "PAFQ_VPA",            "pa_levels",                "Numeric",        "Source",                   "0..*",        "min/day",                                                                                                  "Vigorous physical activity. Key component for muscle mass analysis.",                                                                                                    "—",                                                                                                                        FALSE,
  "CoLaus",   "Physical Activity",          "PAFQ_VPA_pct",        NA,                          "Numeric",        "Source",                   "0..*",        "% of day",                                                                                                 "% daily time in vigorous PA.",                                                                                                                                           "—",                                                                                                                        FALSE,
  "CoLaus",   "Physical Activity",          "pa_levels",           NA,                          "Factor",         "Time-Varying Covariate",   "0..*",        "⚠ OPEN: Levels to be defined",                                                                              "Derived physical activity category from PAFQ components.",                                                                                                               "—",                                                                                                                        TRUE,
  "CoLaus",   "Physical Activity",          "etsem",               NA,                          "Numeric",        "Inactive",                 "0..*",        "kcal/week",                                                                                                "⚠ OPEN: Total weekly energy expenditure — not in dataset. Confirm if required.",                                                                                           "—",                                                                                                                        TRUE,
  "CoLaus",   "Physical Activity",          "etj",                 NA,                          "Numeric",        "Inactive",                 "0..*",        "kcal/day",                                                                                                 "⚠ OPEN: Total daily energy expenditure — not in dataset. Confirm if required.",                                                                                           "—",                                                                                                                        TRUE,
  "CoLaus",   "Physical Activity",          "mnwlk",               NA,                          "Numeric",        "Inactive",                 "0..1",        "min/day",                                                                                                  "Walking time. Baseline only.",                                                                                                                                           "—",                                                                                                                        FALSE,
  "CoLaus",   "Physical Activity",          "phyact",              NA,                          "Factor",         "Inactive",                 "0..*",        "0=Never; 1=Once/wk; 2=Twice/wk; 9=DK",                                                                     "Physical activity frequency (≥20 min/week). Baseline only.",                                                                                                             "—",                                                                                                                        FALSE,

  # ── Dietary — nutrient totals
  "CoLaus",   "Dietary — Nutrient Totals",  "sumtot1",             NA,                          "Numeric",        "Time-Varying Covariate",   "0..*",        "kcal/day",                                                                                                 "Total energy including alcohol. Exclusion threshold: <500 or >4200 kcal/day.",                                                                                           "⚠ OPEN: Confirm thresholds. Carry from adjacent wave if within plausible range.",                                          TRUE,
  "CoLaus",   "Dietary — Nutrient Totals",  "sumtot3",             NA,                          "Numeric",        "Inactive",                 "0..*",        "kcal/day",                                                                                                 "Total energy excluding alcohol.",                                                                                                                                        "—",                                                                                                                        FALSE,
  "CoLaus",   "Dietary — Nutrient Totals",  "sumprot1 / sumprot3", NA,                          "Numeric",        "Inactive",                 "0..*",        "g/day",                                                                                                    "Total protein (incl. / excl. alcohol).",                                                                                                                                 "—",                                                                                                                        FALSE,
  "CoLaus",   "Dietary — Nutrient Totals",  "sumpveg1 / sumpveg3", NA,                          "Numeric",        "Inactive",                 "0..*",        "g/day",                                                                                                    "Vegetal protein (incl. / excl. alcohol).",                                                                                                                                "—",                                                                                                                        FALSE,
  "CoLaus",   "Dietary — Nutrient Totals",  "sumpani1 / sumpani3", NA,                          "Numeric",        "Inactive",                 "0..*",        "g/day",                                                                                                    "Animal protein (incl. / excl. alcohol).",                                                                                                                                 "—",                                                                                                                        FALSE,
  "CoLaus",   "Dietary — Nutrient Totals",  "sumgluc1 / sumgluc3", NA,                          "Numeric",        "Inactive",                 "0..*",        "g/day",                                                                                                    "Total carbohydrates (incl. / excl. alcohol).",                                                                                                                            "—",                                                                                                                        FALSE,
  "CoLaus",   "Dietary — Nutrient Totals",  "sumlipi1 / sumlipi3", NA,                          "Numeric",        "Inactive",                 "0..*",        "g/day",                                                                                                    "Total fat (incl. / excl. alcohol).",                                                                                                                                      "—",                                                                                                                        FALSE,
  "CoLaus",   "Dietary — Nutrient Totals",  "sumvitd1",            NA,                          "Numeric",        "Inactive",                 "0..*",        "µg/day",                                                                                                   "Dietary vitamin D (incl. alcohol).",                                                                                                                                     "—",                                                                                                                        FALSE,
  "CoLaus",   "Dietary — Nutrient Totals",  "sumvitd3",            NA,                          "Numeric",        "Inactive",                 "0..*",        "µg/day",                                                                                                   "Dietary vitamin D (excl. alcohol).",                                                                                                                                     "—",                                                                                                                        FALSE,
  "CoLaus",   "Dietary — Nutrient Totals",  "sumcalc3",            NA,                          "Numeric",        "Inactive",                 "0..*",        "mg/day",                                                                                                   "⚠ OPEN: Dietary calcium — not in current dataset. Confirm if required.",                                                                                                  "—",                                                                                                                        TRUE,
  "CoLaus",   "Dietary — Nutrient Totals",  "pct_prot1",           NA,                          "Numeric",        "Inactive",                 "0..*",        "%",                                                                                                        "% energy from protein (incl. alcohol).",                                                                                                                                 "—",                                                                                                                        FALSE,

  # ── Dairy FFQ amounts
  "CoLaus",   "Dairy — FFQ Items",          "FFQ1amount",          "dairy_total, dairy_fermented, dairy_highfat",  "Numeric (float)", "Source", "0..*", "g/day", "Plain yogurt.", "—", FALSE,
  "CoLaus",   "Dairy — FFQ Items",          "FFQ2amount",          "dairy_total, dairy_fermented, dairy_lowfat",   "Numeric (float)", "Source", "0..*", "g/day", "Low-fat yogurt.", "—", FALSE,
  "CoLaus",   "Dairy — FFQ Items",          "FFQ3amount",          "dairy_total, dairy_fermented, dairy_highfat",  "Numeric (float)", "Source", "0..*", "g/day", "Fruit/aroma yogurt.", "—", FALSE,
  "CoLaus",   "Dairy — FFQ Items",          "FFQ4amount",          "dairy_total, dairy_fermented, dairy_lowfat",   "Numeric (float)", "Source", "0..*", "g/day", "Cottage cheese 0%.", "—", FALSE,
  "CoLaus",   "Dairy — FFQ Items",          "FFQ5amount",          "dairy_total, dairy_fermented, dairy_highfat",  "Numeric (float)", "Source", "0..*", "g/day", "Cottage cheese / ricotta.", "—", FALSE,
  "CoLaus",   "Dairy — FFQ Items",          "FFQ6amount",          "dairy_total, dairy_fermented, dairy_highfat",  "Numeric (float)", "Source", "0..*", "g/day", "Feta / mozzarella.", "—", FALSE,
  "CoLaus",   "Dairy — FFQ Items",          "FFQ7amount",          "dairy_total, dairy_fermented, dairy_highfat",  "Numeric (float)", "Source", "0..*", "g/day", "Gruyère / tomme / camembert.", "—", FALSE,
  "CoLaus",   "Dairy — FFQ Items",          "FFQ8amount",          "dairy_total, dairy_fermented, dairy_highfat",  "Numeric (float)", "Source", "0..*", "g/day", "Cheese fondue.", "—", FALSE,
  "CoLaus",   "Dairy — FFQ Items",          "FFQ52amount",         "dairy_total",               "Numeric (float)", "Source",          "0..*",  "g/day", "Butter.", "—", FALSE,
  "CoLaus",   "Dairy — FFQ Items",          "FFQ53amount",         "dairy_total",               "Numeric (float)", "Source",          "0..*",  "g/day", "Cream 35%.", "—", FALSE,
  "CoLaus",   "Dairy — FFQ Items",          "FFQ63amount",         "dairy_total",               "Numeric (float)", "Source",          "0..*",  "g/day", "Cream tart / cake.", "—", FALSE,
  "CoLaus",   "Dairy — FFQ Items",          "FFQ68amount",         "dairy_total",               "Numeric (float)", "Source",          "0..*",  "g/day", "Ice cream / sorbet.", "—", FALSE,
  "CoLaus",   "Dairy — FFQ Items",          "FFQ71amount",         "dairy_total",               "Numeric (float)", "Source",          "0..*",  "g/day", "Butter for cooking.", "—", FALSE,
  "CoLaus",   "Dairy — FFQ Items",          "FFQ82amount",         "dairy_total, dairy_non_fermented, dairy_lowfat",  "Numeric (float)", "Source", "0..*", "mL/day", "Milk in coffee 0%.", "—", FALSE,
  "CoLaus",   "Dairy — FFQ Items",          "FFQ83amount",         "dairy_total, dairy_non_fermented, dairy_highfat", "Numeric (float)", "Source", "0..*", "mL/day", "Milk in coffee non-0%.", "—", FALSE,
  "CoLaus",   "Dairy — FFQ Items",          "FFQ84amount",         "dairy_total",               "Numeric (float)", "Source",          "0..*",  "mL/day", "Coffee creamer.", "—", FALSE,
  "CoLaus",   "Dairy — FFQ Items",          "FFQ85amount",         "dairy_total, dairy_non_fermented, dairy_lowfat",  "Numeric (float)", "Source", "0..*", "mL/day", "Milk drink 0%.", "—", FALSE,
  "CoLaus",   "Dairy — FFQ Items",          "FFQ86amount",         "dairy_total, dairy_non_fermented, dairy_highfat", "Numeric (float)", "Source", "0..*", "mL/day", "Milk drink non-0%.", "—", FALSE,
  "CoLaus",   "Dairy — FFQ Items",          "FFQ1-86 (freq)",      "FFQfreq",                   "Factor",          "Inactive",        "0..*",  "1=Never … 7=≥2×/day", "Response frequency code. Redundant since amount variables already calculated.", "N/A (inactive)", FALSE,
  "CoLaus",   "Dairy — FFQ Items",          "FFQp1-p86",           "FFQ…amount",               "Factor",          "Inactive",        "0..*",  "1=Less; 2=Equal; 3=More", "Portion size code. Redundant since amount variables already calculated.", "N/A (inactive)", FALSE,
  "CoLaus",   "Dairy — FFQ Items",          "FFQfreq1-86",         "FFQ…amount",               "Numeric (float)", "Inactive",        "0..*",  "—",     "Calculated daily frequency. Redundant since amount variables already calculated.", "N/A (inactive)", FALSE,

  # ── Derived dairy
  "CoLaus",   "Dairy — Derived Variables",  "dairy_total",         NA,                          "Numeric",        "Exposure",                 "0..*",        "g or mL/day",                                                                                              "Sum of all dairy product amounts.",                                                                                                                                      "—",                                                                                                                        FALSE,
  "CoLaus",   "Dairy — Derived Variables",  "dairy_fermented",     NA,                          "Numeric",        "Exposure",                 "0..*",        "g/day",                                                                                                    "Sum of yogurt and cheese amounts. FFQ items: 1, 2, 3, 4, 5, 6, 7, 8.",                                                                                                  "⚠ OPEN: Confirm item list is exhaustive.",                                                                                 TRUE,
  "CoLaus",   "Dairy — Derived Variables",  "dairy_non_fermented", NA,                          "Numeric",        "Exposure",                 "0..*",        "g or mL/day",                                                                                              "Sum of non-fermented dairy amounts. FFQ items: 82, 83, 85, 86.",                                                                                                         "⚠ OPEN: Confirm item list is exhaustive.",                                                                                 TRUE,
  "CoLaus",   "Dairy — Derived Variables",  "dairy_lowfat",        NA,                          "Numeric",        "Exposure",                 "0..*",        "g or mL/day",                                                                                              "Sum of 0%/low-fat dairy. FFQ items: 2, 4, 82, 85.",                                                                                                                     "⚠ OPEN: Confirm item list is exhaustive.",                                                                                 TRUE,
  "CoLaus",   "Dairy — Derived Variables",  "dairy_highfat",       NA,                          "Numeric",        "Exposure",                 "0..*",        "g or mL/day",                                                                                              "Sum of whole-milk, cheese, cream products. FFQ items: 1, 3, 5, 6, 7, 8, 83, 86.",                                                                                       "⚠ OPEN: Confirm item list is exhaustive.",                                                                                 TRUE,
  "CoLaus",   "Dairy — Derived Variables",  "Dairy",               "Dairy_OK",                 "Numeric",        "Source",                   "0..*",        "—",                                                                                                        "Total dairy frequency (excluding butter and cream).",                                                                                                                    "—",                                                                                                                        FALSE,
  "CoLaus",   "Dairy — Derived Variables",  "Dairy_OK",            NA,                          "Binary",         "Inactive",                 "0..*",        "0 = <3/day; 1 = ≥3/day",                                                                                   "Adherence to Swiss dietary guidelines for dairy.",                                                                                                                       "—",                                                                                                                        FALSE,

  # ── OsteoLaus outcomes
  "OsteoLaus","Sarcopenia Outcomes",        "ewgsop2_sarcopenia_stage", NA,                     "Factor",         "Primary Outcome",          "0..*",        "0=None; 1=Probable; 2=Confirmed; 3=Severe",                                                                "EWGSOP2 criteria. Probable: HGS <16 kg (women). Confirmed: Probable + ALM/ht² <5.5 kg/m². Severe: Confirmed + gait speed ≤0.8 m/s.",                                    "Exclude if missing at baseline. ⚠ OPEN: Follow-up missing handling TBD.",                                                  TRUE,
  "OsteoLaus","Sarcopenia Outcomes",        "FNIH_sarcopenia",     NA,                          "Factor",         "Secondary Outcome",        "0..*",        "⚠ OPEN: Levels TBD",                                                                                        "FNIH criteria for sarcopenia sensitivity analysis. ⚠ OPEN: Full definition to be completed.",                               "⚠ OPEN: To be defined.",                                                                                                   TRUE,

  # ── DXA
  "OsteoLaus","DXA — Lean Mass",            "ALM (H_ALM for V5)",  NA,                          "Numeric",        "Primary Outcome",          "0..*",        "kg",                                                                                                       "Appendicular lean mass (arms + legs). V5 variable name: H_ALM.",                                                                                                         "—",                                                                                                                        FALSE,
  "OsteoLaus","DXA — Lean Mass",            "ALM_WT (H_ALM_WT V5)",NA,                          "Numeric",        "Secondary Outcome",        "0..*",        "kg/kg",                                                                                                    "ALM / body weight.",                                                                                                                                                     "—",                                                                                                                        FALSE,
  "OsteoLaus","DXA — Lean Mass",            "ALM_BMI (H_ALM_BMI V5)",NA,                        "Numeric",        "Secondary Outcome",        "0..*",        "kg/m²",                                                                                                    "ALM scaled by BMI.",                                                                                                                                                     "—",                                                                                                                        FALSE,
  "OsteoLaus","DXA — Lean Mass",            "L_…LEAN_MASS / L_ALM_…", NA,                       "—",              "Excluded",                 "0..*",        "—",                                                                                                        "Multiple Lunar DXA lean mass variables (V5). Not used.",                                                                                                                 "N/A (excluded)",                                                                                                           FALSE,

  # ── Gait
  "OsteoLaus","Gait Speed & Functional Tests","6MGS",              "ewgsop2_sarcopenia_stage", "Numeric",        "Primary Outcome",          "1..*",        "m/s",                                                                                                      "6-metre walk test gait speed.",                                                                                                                                          "Exclude if missing.",                                                                                                      FALSE,

  # ── TUG
  "OsteoLaus","Timed Up and Go (V4/V5)",    "TUG_GETUP",           NA,                          "Binary",         "Inactive",                 "0..*",        "⚠ OPEN: 0=?; 1=?",                                                                                          "TUG: Get up from chair with arms crossed.",                                                                                                                              "—",                                                                                                                        TRUE,
  "OsteoLaus","Timed Up and Go (V4/V5)",    "TUG_GO",              NA,                          "Binary",         "Inactive",                 "0..*",        "0=Stop/incomplete; 1=Walk 3m",                                                                             "TUG: Walk 3 metres.",                                                                                                                                                    "—",                                                                                                                        FALSE,
  "OsteoLaus","Timed Up and Go (V4/V5)",    "TUG_TURN",            NA,                          "Binary",         "Inactive",                 "0..*",        "0=Dysbalance/stop/incomplete; 1=Turn back",                                                                "TUG: Turn and return.",                                                                                                                                                  "—",                                                                                                                        FALSE,
  "OsteoLaus","Timed Up and Go (V4/V5)",    "TUG_GOBACKSIT",       NA,                          "Binary",         "Inactive",                 "0..*",        "⚠ OPEN: 0=?; 1=Come back 3m and sit",                                                                       "TUG: Return and sit.",                                                                                                                                                   "—",                                                                                                                        TRUE,
  "OsteoLaus","Timed Up and Go (V4/V5)",    "TUG_TIME",            "TUG_SCORE",                "Numeric",        "Inactive",                 "0..*",        "seconds",                                                                                                  "Time to complete TUG.",                                                                                                                                                  "—",                                                                                                                        FALSE,
  "OsteoLaus","Timed Up and Go (V4/V5)",    "TUG_SCORE",           NA,                          "Factor",         "Inactive",                 "0..*",        "⚠ OPEN: 0–4 levels to be defined",                                                                          "Functional mobility score from TUG time (0–4 points).",                                                                                                                 "—",                                                                                                                        TRUE,
  "OsteoLaus","Timed Up and Go (V4/V5)",    "PHYSTEST_COMMENT",    NA,                          "Character",      "Excluded",                 "0..*",        "—",                                                                                                        "Comments on TUG or gait speed. High missingness; excluded.",                                                                                                             "N/A (excluded)",                                                                                                           FALSE,

  # ── SARC-F
  "OsteoLaus","SARC-F Screening (V5)",      "SARCF_STRENGHT",      NA,                          "Factor",         "Inactive",                 "0..1",        "0=None; 1=Some; 2=A lot or unable",                                                                        "Difficulty lifting/carrying 10 lbs. Note: typo in source name ('STRENGHT').",                                                                                            "—",                                                                                                                        FALSE,
  "OsteoLaus","SARC-F Screening (V5)",      "SARCF_WALK",          NA,                          "Factor",         "Inactive",                 "0..1",        "0=None; 1=Some; 2=A lot/aids/unable",                                                                      "Difficulty walking across a room.",                                                                                                                                      "—",                                                                                                                        FALSE,
  "OsteoLaus","SARC-F Screening (V5)",      "SARCF_CHAIR",         NA,                          "Factor",         "Inactive",                 "0..1",        "0=None; 1=Some; 2=A lot or unable",                                                                        "Difficulty transferring from chair or bed.",                                                                                                                             "—",                                                                                                                        FALSE,
  "OsteoLaus","SARC-F Screening (V5)",      "SARCF_STAIRS",        NA,                          "Factor",         "Inactive",                 "0..1",        "0=None; 1=Some; 2=A lot or unable",                                                                        "Difficulty climbing 10 stairs.",                                                                                                                                         "—",                                                                                                                        FALSE,
  "OsteoLaus","SARC-F Screening (V5)",      "SARCF_FALL",          NA,                          "Factor",         "Inactive",                 "0..1",        "0=None; 1=1–3 falls; 2=≥4 falls",                                                                          "Falls in the past year.",                                                                                                                                                "—",                                                                                                                        FALSE,
  "OsteoLaus","SARC-F Screening (V5)",      "SARCF_TOTAL",         NA,                          "Integer",        "Inactive",                 "0..1",        "0–10",                                                                                                     "SARC-F total score.",                                                                                                                                                    "—",                                                                                                                        FALSE,

  # ── HGS OsteoLaus
  "OsteoLaus","Handgrip Strength (V5)",     "HGS_R1, HGS_R2, HGS_R3", "HGS_MAX",              "Numeric",        "Source",                   "0..1",        "kg",                                                                                                       "Three right-hand grip measures. Recorded if right hand is dominant.",                                                                                                    "—",                                                                                                                        FALSE,
  "OsteoLaus","Handgrip Strength (V5)",     "HGS_L1, HGS_L2, HGS_L3", "HGS_MAX",              "Numeric",        "Source",                   "0..1",        "kg",                                                                                                       "Three left-hand grip measures. Recorded if left hand is dominant.",                                                                                                      "—",                                                                                                                        FALSE,
  "OsteoLaus","Handgrip Strength (V5)",     "HGS_MAX",             NA,                          "Numeric",        "Secondary Outcome",        "0..*",        "kg",                                                                                                       "Maximum grip strength across all 6 trials.",                                                                                                                             "—",                                                                                                                        FALSE,

  # ── Other
  "OsteoLaus","Other",                      "Death",               NA,                          "Factor",         "Sensitivity Analysis",     "0..*",        "—",                                                                                                        "⚠ OPEN: Visit at which death occurred. Used to exclude participants dying within 1–2 years of last scan (cachexia). V5 only. Not in current dataset — confirm if required.", "—",                                                                                                                       TRUE,
)
```

```{r}
#| label: open-questions-count
open_n <- sum(vars$open)
```

```{r}
#| label: full-table

vars |>
  mutate(
    original   = paste0("`", original, "`"),
    derived    = if_else(is.na(derived), "—", paste0("`", derived, "`")),
    open_flag  = if_else(open, "⚠", "")
  ) |>
  select(
    `⚠`       = open_flag,
    Source     = source,
    Section    = section,
    `Variable` = original,
    `Derived`  = derived,
    Type       = data_type,
    Role       = role,
    Mult.      = multiplicity,
    Levels     = levels,
    Notes      = notes,
    `Missing handling` = missing
  ) |>
  datatable(
    filter    = "top",
    rownames  = FALSE,
    escape    = FALSE,
    options   = list(
      pageLength  = 20,
      scrollX     = TRUE,
      autoWidth   = FALSE,
      columnDefs  = list(
        list(width = "60px",  targets = 0),
        list(width = "80px",  targets = 1),
        list(width = "160px", targets = 2),
        list(width = "120px", targets = 3),
        list(width = "120px", targets = 4),
        list(width = "100px", targets = 5),
        list(width = "140px", targets = 6),
        list(width = "60px",  targets = 7),
        list(width = "160px", targets = 8),
        list(width = "260px", targets = 9),
        list(width = "200px", targets = 10)
      )
    ),
    class = "compact stripe hover"
  ) |>
  formatStyle(
    "⚠",
    color            = "red",
    fontWeight        = "bold"
  ) |>
  formatStyle(
    "Role",
    backgroundColor = styleEqual(
      names(ROLE_COLOURS),
      unname(ROLE_COLOURS)
    ),
    color = styleEqual(
      names(ROLE_COLOURS),
      ifelse(names(ROLE_COLOURS) %in% c("Inactive", "Excluded", "Source"), "black", "white")
    )
  )
```

------------------------------------------------------------------------

# Variable Derivation Logic {#sec-derivation}

This section documents how each derived analytical variable is constructed from its source variables. These rules are the authoritative reference for the data cleaning script.

## `alcohol_category`

`alcohol_category` is a four-level time-varying factor (1 = Non-drinker, 2 = Light, 3 = Moderate, 4 = Heavy) derived from two source variables that differ in availability across waves.

At **Baseline**, `sumalco` (grams of alcohol per day) is missing by design. `alcohol_category` is therefore derived exclusively from `conso_hebdo` (alcohol in units per week), converting units to grams per day using a factor of 10–12 g per unit. From **F1 onwards**, `sumalco` is the preferred source; `conso_hebdo` is used as a fallback if `sumalco` is missing at a given wave. If both sources are missing at a wave, the category is carried forward from the most recent available wave or, failing that, backward from the nearest future wave. A participant is excluded if no category can be imputed at any wave.

The two validation variables `alcuse` (binary: any alcohol use) and `alcool4` (binary: drinker/non-drinker) are used to cross-check the non-drinker classification and should be compared for data completeness.

## `diabetes_status`

`diabetes_status` is a four-level time-varying factor (1 = No diabetes, 2 = Diabetes treated with oral agents or diagnosed by laboratory criteria only, 3 = Diabetes requiring insulin, 4 = Impaired glucose tolerance / prediabetes) derived from eight source variables.

The hierarchy of evidence is: laboratory markers (`DIAB`: fasting plasma glucose ≥ 7.0 mmol/L; `DIAB_Hb`: HbA1c ≥ 48 mmol/mol, available from F2 onwards) take precedence over self-report (`dbtld`: doctor-diagnosed diabetes; `dbdrg`: treatment history). Medication flags (`insn`: insulin use; `orldrg`: oral antidiabetics; `antiDIAB`: any antidiabetic drug) determine whether Level 2 (oral/lab only) is upgraded to Level 3 (insulin-requiring). `DIAB2` (combined IGT/prediabetes flag) is used as a validation variable to identify Level 4 participants, though its coding differs across waves (three levels at Baseline/F2/F3; binary at F1 — see open question).

## `cdv_event`

`cdv_event` is a binary time-varying composite cardiovascular event indicator (0 = No, 1 = Yes) derived from thirteen individual self-reported cardiovascular history variables: myocardial infarction (`miac`), stroke (`strk`), congestive heart failure (`chf`), coronary artery disease (`cad`), angina (`angn`), cardiomyopathy (`cmp`), hypertensive heart disease (`hdc`), heart valve disease (`hdv`), arrhythmia (`artm`), vascular surgery (`vslg`), cardiac catheterisation (`ccth`), bypass surgery/CABG (`cabg`), and stent/PCI (`pcin`).

`cdv_event` is coded 1 if at least one component variable is positive (value = 1). It is coded 0 if all component variables are 0. A participant is excluded if all component variables are missing at a given wave. The value 9 ("does not know") in individual components is treated as missing for the purpose of deriving `cdv_event`.

## `ewgsop2_sarcopenia_stage`

`ewgsop2_sarcopenia_stage` is a four-level ordinal outcome (0 = No sarcopenia, 1 = Probable, 2 = Confirmed, 3 = Severe) following the EWGSOP2 2019 consensus criteria, applied to women only.

**Probable sarcopenia** (Level 1) is assigned when handgrip strength (`handgrip` in CoLaus waves; `HGS_MAX` in OsteoLaus V5) is below 16 kg. **Confirmed sarcopenia** (Level 2) requires Level 1 criteria plus low muscle quantity, defined as appendicular lean mass indexed to height squared (ALM / ht²) below 5.5 kg/m² from DXA. **Severe sarcopenia** (Level 3) requires Level 2 criteria plus impaired physical performance, defined as 6-metre gait speed (`6MGS`) at or below 0.8 m/s. A participant is excluded from the primary analysis if `ewgsop2_sarcopenia_stage` is missing at baseline.

## `hrt_status`

`hrt_status` is a three-level time-varying factor (Never, Former, Current) for hormone replacement therapy use, derived from five source variables.

The Baseline value is back-filled using information collected at F1 and later waves. `esthrp` (ever taken HRT) and `esthrpage` (age at HRT start) are the primary Baseline screening variables; they are used together to determine whether HRT was active at the time of the Baseline examination. From F1 onwards, `bthc` (current HRT use at time of visit) provides the current status directly. `subsp` (age at cessation) and `subyr` (cumulative exposure duration) are used to classify participants as Former users and to characterise the duration of past exposure.

## Dairy exposure variables

Five dairy exposure variables are computed by summing the daily amount variables (in grams or millilitres per day) from the FFQ. All FFQ amount variables are available from F1 onwards.

`dairy_total` is the sum of all dairy product amounts across all relevant FFQ items. `dairy_fermented` sums yogurt and cheese items (FFQ items 1–8). `dairy_non_fermented` sums milk-based drink and coffee milk items (FFQ items 82, 83, 85, 86). `dairy_lowfat` sums 0% fat products (FFQ items 2, 4, 82, 85). `dairy_highfat` sums full-fat products including whole-milk cheeses, cream, and full-fat milks (FFQ items 1, 3, 5, 6, 7, 8, 83, 86). The item lists for each sub-category should be verified to confirm no items are missing or double-counted (see open questions).

The raw frequency codes (`FFQ1–86`), portion size codes (`FFQp1–p86`), and calculated daily frequency variables (`FFQfreq1–86`) are redundant given that the amount variables already integrate frequency and portion size, and are therefore marked Inactive.

------------------------------------------------------------------------

# Open Questions {#sec-open-questions}

The following items require investigator decision. This section is auto-generated from the `open` flag in the variable table.

```{r}
#| label: open-questions-table

vars |>
  filter(open) |>
  mutate(item = row_number()) |>
  select(
    `#`      = item,
    Source   = source,
    Section  = section,
    Variable = original,
    `Open question (from Notes / Missing columns)` = notes
  ) |>
  kable(format = "html") |>
  kableExtra::kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"),
    full_width = TRUE
  ) |>
  kableExtra::column_spec(4, monospace = TRUE, bold = TRUE) |>
  kableExtra::row_spec(0, bold = TRUE, background = "#E84855", color = "white")
```

::: callout-important
## Action required

Each row above must be resolved by the responsible investigator. Once resolved:

1.  Update the `notes` and `missing` fields in the `vars` tibble in this document.
2.  Set `open = FALSE` for that variable.
3.  Re-render this document (`quarto render variable_overview.qmd`).
4.  Commit the updated `.qmd` to version control with a descriptive message.
:::

------------------------------------------------------------------------

# Session Information {#sec-session}

Rendered for full computational reproducibility. This block records the exact R environment used.

```{r}
#| label: session-info

sessionInfo()
```

```{r}
#| label: render-timestamp
#| echo: false
cat("Document rendered:", format(Sys.time(), "%Y-%m-%d %H:%M:%S %Z"))
```
